---
title: "Tour Reporting"
author: "KGW"
date: "February 12, 2016"
output: html_document
---

We're going to quietly obtain the data here.  It's being drawn from JSON files pulled from TrainerRoad over the course of the Tour and then tidied and summarized for reporting.

```{r, echo=FALSE, message=F, warning=F, cache=FALSE}

library(jsonlite)
library(lubridate)
library(dplyr)
library(tidyr)
library(knitr)
library(ggplot2)
library(car)
library(xts)
library(dygraphs)
library(RColorBrewer)

.pardefault <- par()

#data-obtaining, given manual file munging
#highest ID: 3185481
setwd("~/datacourse/small projects/tour_data")

filepath <- "lists-of-rides/trall/"

#load our reference tables
tzones <- read.csv("timezones.csv", stringsAsFactors = FALSE)
names(tzones) <- c("TimeZoneId", "OlsonTime", "Offset","Order")

tzones$name <- ifelse(tzones$Offset > 0, 
       paste("+",as.character(tzones$Offset),sep = ""),           
       as.character(tzones$Offset))

stages <- read.csv("stagemap.csv",stringsAsFactors = FALSE)
meters <- read.csv("powersources.csv", stringsAsFactors = FALSE)

d <- NULL

file.names <- dir(filepath, pattern = ".json")
#start to build the file
file <- fromJSON(paste(filepath,file.names[1],sep = ''))
file <- file$Rides[,1:21]
d <- file
#add on new records
for (i in 2:length(file.names)) {
     file <- fromJSON(paste(filepath,file.names[i],sep = ''))
     rides <- file$Rides[,1:21]
     d <- rbind(d, rides)
}

rm(file)
rm(rides)

#clean out duplicates and sort
d <- d %>% distinct() %>% arrange(desc(Id))

#map in the stages
d <- left_join(d, stages, by="WorkoutName")

#drop the Not ToS stuff
d <- filter(d, substring(Stage,1,5) == "Stage")

#add time zones
d <- left_join(d,tzones,by = "TimeZoneId")

#sort out the time zones - that loop is ugly but will do for now
d$pTime <- ymd_hms(d$WorkoutDate)

for (i in c(1:nrow(d))) {
     d[i,"pTime"] <- ymd_hms(d[i,"WorkoutDate"], tz = d[i,"OlsonTime"])
}

#drop the early starts
d <- filter(d, d$pTime >= ymd_hms("2016-02-04 10:00:00", tz = "UTC"))

#go get the trainers
d <- left_join(d, meters, by = "PowerSource")

#compile rider performance
d$RiderTSSRatio <- d$TSS/d$WorkoutTSS
d$RiderIFRatio <- d$IntensityFactor/d$WorkoutIF

#drop the crazy outliers
qnt <- quantile(d$RiderTSSRatio, probs = c(.001, .999))
d <- d[d$RiderTSSRatio < qnt[2],]
d <- d[d$RiderTSSRatio > qnt[1],]

qnt <- quantile(d$Lthr, probs = c(.001, .999))
d <- d[d$Lthr < qnt[2],]
d <- d[d$Lthr > qnt[1],]

qnt <- quantile(d$Ftp, probs = c(.001, .999))
d <- d[d$Ftp < qnt[2],]
d <- d[d$Ftp > qnt[1],]

#rider summary table
riders <- d %>% 
     group_by(MemberId) %>%
     summarize("Workouts"=n(), 
               "FTP" = median(Ftp),
               "Stages Ridden"= n_distinct(Stage),
               "PowerSourceType"=first(SourceType),
               "LTHR" = median(Lthr)
     )

#hourly summary table
hourly <- d %>%
     group_by(Stage, "Time"=round_date(pTime, unit = "hour")) %>%
     summarise("HourCount"=n()) %>%
     mutate("Cumulative"=cumsum(HourCount)) %>%
     arrange(Time)

hourly <- as.data.frame(hourly)

# build cumulative and spot hourly stage counts
# getting the leading zeroes out before filling in gaps
# so as to avoid spiky graphs.

# This Bunch are cumulative

yCumulatives <- spread(hourly[,c(1,2,4)], Stage, Cumulative)
for(i in c(2:10)) {
     j <- 1
     while(is.na(yCumulatives[j,i])) {
          yCumulatives[j,i] <- 0
          j <- j+1
     }
}
for(i in c(2:10)) {yCumulatives[,i] <- na.locf(yCumulatives[,i], na.rm = FALSE)}

#long form data better for ggplot2
yCumulativesLong <- gather(yCumulatives,"Stage", "Count", 2:10)

# now the raw counts - we'll pull the counts; 
# then smooth them with loess().  And make a long one.
# for this one, we lose leading and trailing zeroes and then fill.
yCounts <- spread(hourly[,c(1,2,3)], Stage, HourCount)
for(i in c(2:10)) {
     j <- 1
     while(is.na(yCounts[j,i])) {
          yCounts[j,i] <- 0
          j <- j+1
     }
     j <- nrow(yCounts)
     while(is.na(yCounts[j,i])) {
          yCounts[j,i] <- 0
          j <- j-1
     }
}

for(i in c(2:10)) {yCounts[,i] <- na.locf(yCounts[,i], na.rm = FALSE)}

yCountsLong <- gather(yCounts,"Stage", "Count", 2:10)

# smooooth.
for(i in c(2:10)) {
     lo <- loess(as.formula(paste("yCounts$`",names(yCounts)[i],"`","~ as.numeric(Time)", sep="")), 
                 data = yCounts, span = 0.25)
     yCounts[,i+9] <- predict(lo)
     names(yCounts)[i+9] <- paste("Smooth",i-1)
}

yCountsSmooth <- select(yCounts, Time, starts_with("Smooth"))
names(yCountsSmooth) <- gsub("Smooth", "Stage", names(yCountsSmooth))
yCountsSmoothLong <- gather(yCountsSmooth, "Stage", "Count", 2:10)
yCountsSmoothLong[yCountsSmoothLong$Count < 0, "Count"] <- 0


```

_Latest Ride Recorded_: `r max(d$pTime)` UTC

_Latest Ride ID_: `r max(d$Id)`

How have rides built over time?

```{r}
t <- ggplot(yCumulativesLong, aes(Time, Count, fill = Stage)) +
     geom_area()
t
```


What is the flow of rides from hour to hour?

```{r fig.width=10}

# v <- ggplot(yCountsSmoothLong, aes(Time, Count, fill = Stage)) +
#      geom_area() + 
#      ggtitle("Stages Evolving in Real Time (UTC)")
# v

dygraph(xts(yCountsSmooth[,-1], order.by = yCountsSmooth$Time),
        main = "Rides per Hour per Stage (UTC)") %>%
     dyOptions(stackedGraph = TRUE, useDataTimezone = TRUE,
               colors = RColorBrewer::brewer.pal(8, "Dark2")) 
```

How many rides have been recorded per stage so far?

```{r warning=FALSE, fig.width = 7}
q <- ggplot(d, aes(Stage))
q + geom_bar()

```

So, we've calculated the ratio between the Stage's expected TSS and the rider's actual TSS for that ride.  1.0 means they exerted themselves exactly as much as expected by TrainerRoad

```{r}
r <- ggplot(d, aes(Stage, RiderTSSRatio))
r + geom_boxplot() + ggtitle("Riders' TSS vs Expected - 1.0 is equal")

```

OK.  What about IF?  Is that sinking as we go along?

```{r}
r <- ggplot(d, aes(Stage, RiderIFRatio))
r + geom_boxplot() + ggtitle("Riders' IF vs Expected - 1.0 is equal")

```

```{r fig.height=7, fig.width=7, echo=FALSE, warning=FALSE}
par(.pardefault)
```


Simple matrix comparisons:

```{r, echo=FALSE, fig.width=10, fig.height=10}

scatterplotMatrix(~TotalTicks+NormalizedPower+TSS+IntensityFactor+Ftp|Stage, 
                  data=d, main="Stages")
```

Let's see some simple statistics:
```{r stats, fig.width=7, fig.height=7}
sumtabl <- d %>% 
     group_by(Stage, Sufferfest) %>% 
     summarise(n(), 'Athletes' = n_distinct(MemberId), 'Total Hours'= (sum(TotalTicks)/3600), 'Total TSS' = sum(TSS), 'Avg NP' = mean(NormalizedPower), 'Median NP' = median(NormalizedPower), 'Avg IF' = mean(IntensityFactor), 'Median IF' = median(IntensityFactor))

kable(sumtabl, digits = 2)
```

So who are these people? (Total number of riders reporting: *`r nrow(riders)`*)

```{r fig.width=10}
par(mfrow = c(1,2))
hist(riders$FTP,breaks = 20)
hist(riders$LTHR,breaks = 30)
par(mfrow=c(1,1))
```

```{r meters, echo=FALSE}
# What kind of power do various meter users report?

# pmtabl <- d %>% group_by(PowerSource) %>% summarise(n(), "Reported FTP"=median(Ftp), "Median Ride NP" = median(NormalizedPower))
# 
# kable(pmtabl, digits = 2)
```

Do riders using power meters report different FTP or NP/ride than those using virtual power?

```{r}
f <- ggplot(d, aes(Ftp, NormalizedPower, color = SourceType))
f <- f + 
     geom_point() + 
     xlab("Rider Reported FTP") + 
     ylab("Ride NP") + 
     geom_segment(x=0,xend=400,y=0,yend=400)
f

```

What time of day are people riding? We've dropped the UTC folks because we really don't know whether they're British or their systems report in UTC. (Hmmm. Do these people not have jobs?)

```{r}
hist(hour(ymd_hms(d["OlsonTime" != "UTC", "WorkoutDate"])), 
     xlab = "Hour of Day",
     main = "Hour Rides Began (UTC Times Dropped)")
```
